Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
beta gone

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model Algorithm}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} \manuel{$\rho$: learning rate,}{$r_1, r_2$: parameters controlling the learning rate} 
    % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign \{\}$
    \For{$i \assign 0$ \textbf{to} $\minit$}
    \State $\sigma \assign \text{generate uniformly at random}$
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor
    %\State $S \assign $ generate  $\minit$ permutations uniformly at random
    %\State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}\Comment{Learning Step}
    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}
    \Statex \Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
    \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    % \Comment{Generate $\beta$ permutations}
    \State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
    \Statex \Comment{Select the farthest from $S$}
   % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor 
    \State \Return $\tuple{\sigma^\ast, f(\sigma^\ast)}\assign \argmin_{\tuple{\sigma, f(\sigma)}\in S} f(\sigma)$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

@@

To increase the exploration of the search space, the algorithm samples $\beta$ permutations, $\Sigma$. Then it selects  for evaluation the farthest from permutations already evaluated ($S$), as shown in line~\ref{line:distance} of Algorithm~\ref{alg:umm}.
%i.e.,  $\sigma\in \Sigma$ such that $\sigma = \argmax_{\sigma'\in \Sigma}\min d(\sigma', \pi)$ for every $\pi\in S$.

\EKHINE{The algorithm uses a parameter $\beta$ that determines the number of points sampled, while the authors analyze the sensitivity of the results to the parameters r 1 and r 2, they do not do that with $\beta$. In my opinion the value of $\beta$ is quite relevant.}

@@

\subsubsection{UMM}
%
UMM uses the same budget of evaluations ($\FEmax$) as CEGO and initially
samples $\minit=10$ solutions uniformly.  At each sampling step, we set
$\beta=10$, meaning that 10 permutations are generated and the farthest from
$S$ is selected for evaluation.  The other two parameters of UMM are $r_1$ and
$r_2$, which determine the balance between exploration and exploitation. In the
experimental section we will study their impact.

@@

\ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
 obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
instance considered here. The lack of any apparent convergence suggests that
the models (both Bayesian and probabilistic) are not learning anything about
the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{kaka}
% \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
% that other distance measures are more suited than Kendall's-$\tau$ distance for
% the PFSP, the reported fitness differences between various distance metrics are
% small.
% Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}

>>>>

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model Algorithm}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $\FEmax$: total budget,\newline\hspace*{5ex} \manuel{$\rho$: learning rate,}{$r_1, r_2$: parameters controlling the learning rate} 
    % \\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign \{\}$
    \For{$i \assign 0$ \textbf{to} $\minit$}
    \State $\sigma \assign \text{generate uniformly at random}$
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor
    %\State $S \assign $ generate  $\minit$ permutations uniformly at random
    %\State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}\Comment{Learning Step}
    \State Initialize (first iteration) or decrease  $\theta$\label{line:theta}
    \Statex \Comment{See \emph{Sampling Step} and Eq.~\ref{eq:expectation}}
   % \State $\Sigma \assign$ sample $\beta$ permutations from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
     \State $\sigma \assign$ sample a permutation from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    % \Comment{Generate $\beta$ permutations}
    %\State $\sigma\assign  \argmax_{\pi'\in \Sigma}\min_{ \pi \in S} d(\pi', \pi)   $ \label{line:distance}
    %\Statex \Comment{Select the farthest from $S$}
   % \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \tuple{\sigma, f(\sigma)}$\Comment{Evaluate it}
    \EndFor 
    \State \Return $\tuple{\sigma^\ast, f(\sigma^\ast)}\assign \argmin_{\tuple{\sigma, f(\sigma)}\in S} f(\sigma)$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

@@

% To increase the exploration of the search space, the algorithm samples $\beta$ permutations, $\Sigma$. Then it selects  for evaluation the farthest from permutations already evaluated ($S$), as shown in line~\ref{line:distance} of Algorithm~\ref{alg:umm}.
% %i.e.,  $\sigma\in \Sigma$ such that $\sigma = \argmax_{\sigma'\in \Sigma}\min d(\sigma', \pi)$ for every $\pi\in S$.

% \EKHINE{The algorithm uses a parameter $\beta$ that determines the number of points sampled, while the authors analyze the sensitivity of the results to the parameters r 1 and r 2, they do not do that with $\beta$. In my opinion the value of $\beta$ is quite relevant.}

@@

\subsubsection{UMM}
%
UMM uses the same budget of evaluations ($\FEmax$) as CEGO and initially
samples $\minit=10$ solutions uniformly.
%At each sampling step, we set
%$\beta=10$, meaning that 10 permutations are generated and the farthest from
%$S$ is selected for evaluation.
The other two parameters of UMM are $r_1$ and $r_2$, which determine the
balance between exploration and exploitation. In the experimental section we
will study their impact.

@@

% \ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
%  obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
% instance considered here. The lack of any apparent convergence suggests that
% the models (both Bayesian and probabilistic) are not learning anything about
% the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{}
% \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
% that other distance measures are more suited than Kendall's-$\tau$ distance for
% the PFSP, the reported fitness differences between various distance metrics are
% small.
% Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}
